---
title: "Opportunity Youth Population Estimates and Demographics (Preliminary)"
author: "Diego Rubi"
date: "9/22/2020"
output: powerpoint_presentation
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = F, 
                      message = F, 
                      fig.height = 4, 
                      fig.width = 8)
```

```{r load-libraries-functions}

source(
  here::here(
    'R', 
    'presentation_setup.R'
  )
)


```

```{r load-data}

analysis_data <- 
  load_data$load_population_estimates()[['q1_demographics']]

```


```{r formatting, include = F}


analysis_data <- 
  imap(
    .x = analysis_data, 
    .f = ~ 
      if(.y %in% c("total_population_age_bracket", 
                   'geo_puma_pop')){
        .x 
      }else{
        .x %>% 
          mutate(
            oy_flag = figure_helpers$format_oy_variable(oy_flag)
          ) %>% 
          filter(oy_flag != "Everyone Else")
          
      }
  )

analysis_data$q2_household <- 
  modify(.x = analysis_data$q2_household, 
         .f = ~.x %>% mutate(oy_flag = figure_helpers$format_oy_variable(oy_flag)))

erase_axis = 
   theme(
    axis.text.y = element_blank(), 
    axis.line.y = element_blank(), 
    axis.title.y = element_blank(), 
    axis.ticks.y = element_blank()
  )


# Settings for the lower error bar
# other option is percent_low
error_bar_low <- 
  rlang::sym('percent')

# color contrasts between Connected Youth and Opportunity Youth
# color_contrast <- 
#   c(`Connected Youth` = "#800000", `Opportunity Youth` = '#D5802B')

color_contrast <- 
  c(`Connected Youth` = "#767676", `Opportunity Youth` = '#D5802B')

```




## Today's Agenda

- Population of Chicago by Age Bracket 
- Total Number of Opportunity Youth in Chicago
- Race and Ethnicity Distribution
- Gender distribution
- Disability distribution
- Age Bracket
- Pending analyses: 
  - Race & Gender
  - Population by PUMA

## Methodology

- 2018 American Community Survey Public Use Microdata
  - Limited to Chicago relevant Public Use Microdata Areas (PUMA)
- Opportunity Youth are individuals who respond as follows to questions about school enrollment, employment, and age: 
  - AGE: Between 16 - 24
  - School Enrollment: Have not attended school within the last three months
  - Employment: Are 'unemployed' or 'not in the labor force'

## About 12% of Chicago's population is within 16 and 24 years of age

```{r}
age_population <- analysis_data$total_population_age_bracket %>% select(age_bracket, n) %>% spread(key = age_bracket, value = n) %>% unlist()
```


```{r}


analysis_data$total_population_age_bracket %>% 
  
  ggplot(aes(x = age_bracket, y = percent)) + 
  geom_col(width = .85, 
           color = 'black',
           fill = "#D5802B") + 
  geom_text(aes(label = 
                    scales::percent(percent)), 
            vjust = -.25, 
            position = position_stack(vjust = .5)) + 
  geom_errorbar(
    aes(ymin = !!error_bar_low, 
        ymax = percent_upp), 
    width = .1
  ) + 
theme_classic() + 
  labs(x = "Age bracket of Chicago Population", 
       y = "Percent") + 
  scale_y_continuous(labels = scales::percent) + 
  erase_axis + 
  scale_x_discrete(labels = function(x) paste0(x, 
                                               figure_helpers$fig_help$newline, 
                                               figure_helpers$fig_help$N,
                                               format(age_population, big.mark = ',')))
 
```

```{r}
total_sample <- 
  
  analysis_data$total_oy_population %>% 
  select(oy_flag, n) %>% 
  arrange(oy_flag) %>% 
  pivot_wider(names_from = oy_flag, values_from = n) %>% 
  unlist()
```


## About `r (total_sample/sum(total_sample))[2] %>% scales::percent()` of 16 to 24-year olds are Opportunity Youth

```{r}



analysis_data$total_oy_population %>% 
  # filter(oy_flag != "Everyone Else") %>% 
  
  ggplot(aes(x = oy_flag, 
             y = percent)) + 
  geom_col(width = .5, 
           fill = color_contrast,
           color = 'black') + 
  geom_text(aes(label = scales::percent(percent, accuracy = .1)),
            position = position_stack(.5))+
  geom_errorbar(
    mapping = aes(ymin = !!error_bar_low, ymax = percent_upp), 
    width = .1
  ) + 
  theme_classic() + 

  scale_x_discrete(labels = function(x) paste0(x, 
                                               figure_helpers$fig_help$newline,
                                               figure_helpers$fig_help$N,
                                                 format(total_sample, big.mark = ',')))+ 
  labs(
    x = "",
    y = 'Population Estimate', 
    caption = 
      str_wrap("*A subset of the population of 16-24 year olds is not captured as either Opportunity Youth or Connected Youth due to missing responses in the data. ", 
               width = 80)
  ) + 
  scale_y_continuous(labels = scales::percent) + 
  erase_axis
```




## More than half of all Opportunity Youth are Black, Non-Hispanic 

```{r race, echo = F, eval = F}
analysis_data$oy_race %>% 
  ggplot(aes(x = oy_flag, 
         y = percent)) + 
  geom_col(
    aes(fill = race_ethnicity)
  ) + 
  geom_text(aes(
    label = if_else(percent >= .08, 
                    scales::percent(percent, accuracy = .1), 
                    ""), 
    group = race_ethnicity), 
    position = position_stack(vjust = .5)
  ) + 
  labs(x = figure_helpers$x_axis, 
       y = figure_helpers$y_percent, 
       fill = "Race / Ethnicity") + 
  scale_y_continuous(labels = scales::percent) + 
  theme_classic() + 
  theme(
    legend.position = 'right'
  ) + 
  scale_fill_discrete(labels = function(x) str_wrap(x, 15)) + 
  scale_fill_brewer(direction = -1) + 
  erase_axis

       
```

```{r race-alterante}

analysis_data$oy_race_alternate %>% 
  ggplot(aes(x = oy_flag, y = percent)) + 
  geom_col(aes(fill = race_alternate), 
           width = .5)+ 
  geom_text(aes(label =
                  if_else(percent > .05, 
                          scales::percent(percent, accuracy = .1),
                          ""),
                group = race_alternate), 
            position = position_stack(vjust = .5)) + 
  theme_classic() + 
  labs(
    y = figure_helpers$y_percent, 
    x = "",
    fill = "Race or Ethnicity", 
    caption = 
      "* Percentages less than 5% suppressed for clarity"
    
  ) + 
  ggsci::scale_fill_uchicago() + 
  # scale_fill_brewer(palette = 7, direction = 1, type = 'seq') + 
  erase_axis
```


## Race / Ethnicity Alternate Visualization

```{r}


analysis_data$oy_race_alternate %>% 
  ggplot(aes(x = oy_flag, 
         y = percent)) + 
  geom_col(aes(fill = race_alternate), 
           position = 'dodge', 
           color = 'black') + 
  geom_errorbar(
    aes(ymin = !!error_bar_low, 
        ymax = percent_upp, 
        group = race_alternate), 
    width = .1,
    position = position_dodge(width = .9, preserve = 'total')
  )+ 
  geom_text(aes(
    label = if_else(percent >= .05,
                    scales::percent(percent, accuracy = .1),
                    ""),
    group = race_alternate),
    position = position_dodge(width = .9), 
    vjust = 2, 
    size = 3.2
  ) +
  labs(x = "", 
       y = figure_helpers$y_percent, 
       fill = "Race / Ethnicity", 
       caption = "* Percentages less than 5% suppressed for clarity"
       ) + 
  scale_y_continuous(labels = scales::percent) + 
  theme_classic() + 
  theme(
    legend.position = 'right'
  ) + 
  scale_fill_discrete(labels = function(x) str_wrap(x, 15)) + 
  # scale_fill_brewer(palette = 7, direction = 1) 
  ggsci::scale_fill_uchicago()



```


## Nearly 60% of Opportunity Youth identify as male 

```{r sex}

# Male

analysis_data$oy_sex %>% 
  filter(SEX_label != "Female") %>% 
  ggplot(aes(x = oy_flag, 
             y = percent)) + 
  geom_col(width = .5, 
           # color = 'black',
           aes(fill = oy_flag), 
           color = 'black'
  ) + 
  geom_errorbar( aes(ymax = percent_upp ,
                     ymin = !!error_bar_low), 
                 width = .1) + 
  geom_text(
    aes(
      label = scales::percent(percent, accuracy = .1), 
      group = SEX_label), 
      position = position_stack(vjust = .5), 
    vjust = -.25
    ) + 
  labs(x = "Percentage of individuals between 16-24 who identify as male", 
       y = figure_helpers$y_percent, 
       fill = "Sex / Gender") + 
  scale_y_continuous(labels = scales::percent) + 
  theme_classic() + 
  theme(
    legend.position = 'bottom'
  ) + 
  erase_axis  +
  scale_fill_manual(values = c(color_contrast)) + 
  guides(fill = FALSE)
```


```{r disability}

analysis_data$oy_disability %>% 
  filter(DIS_label == "With a disability") %>%
  ggplot(aes(x = oy_flag,y = percent)) + 
  geom_col(aes(fill = oy_flag),
           width = .5, 
           color = 'black') + 
  geom_text(
    aes(
      label = scales::percent(percent, accuracy = .1), 
      group = DIS_label), 
    position = position_stack(vjust = .5), 
    vjust = -.25
  ) + 
  geom_errorbar(mapping = aes(ymin = !!error_bar_low, ymax = percent_upp), 
                width = .1) +
  labs(x =  "Percentage of individuals between 16-24 that report having a disability", 
       y = figure_helpers$y_percent, 
       color = "Disability Status") + 
  scale_y_continuous(labels = scales::percent) + 
  theme_classic() + 
  theme(
    legend.position = 'bottom'
  ) + 
  erase_axis + 
  scale_fill_manual(values = color_contrast) + 
  guides(fill = FALSE)
  
```

## Opportunity Youth are more likely to be older


```{r age-brackets}
analysis_data$oy_age_bracket %>% 
  filter(oy_flag != "Everyone Else", 
         age_bracket != "16-19") %>% 
  ggplot(aes(x = oy_flag, 
         y = percent)) + 
  geom_col(
    # aes(fill = age_bracket),
    fill = color_contrast,
    color = 'black', 
    width = .5
  ) + 
  geom_text(
    aes(
      label = scales::percent(percent, accuracy = .1), 
      group = age_bracket), 
      position = position_stack(vjust = .5), 
    vjust = -.25
    ) + 
  geom_errorbar(
    mapping = aes(ymin = !!error_bar_low, ymax = percent_upp), 
    width = .1
  ) +
  labs(x = "Percentage of individuals between 20-24", 
       y = figure_helpers$y_percent, 
       fill = "Age Bracket") + 
  scale_y_continuous(labels = scales::percent) + 
  theme_classic() + 
  theme(
    legend.position = 'bottom'
  ) +
  
  erase_axis + 
  scale_fill_brewer(palette = 7)


```

## Educational Attainment

```{r}
analysis_data$oy_edu_attainment %>% 
  filter(oy_flag != "Everyone Else") %>% 
  ggplot(aes(x = oy_flag, y = percent)) + 
  geom_col(aes(fill = SCHL_label)) + 
  theme_classic() + 
  labs(
    fill = "Educational Attainment", 
    y = "Percent"
  ) + 
  erase_axis + 
  theme(axis.title.x = element_blank()) + 
  geom_text(aes(label = if_else(percent > .05, 
                                scales::percent(percent, accuracy = .1), 
                                ""), 
                group = SCHL_label), 
            position = position_stack(vjust = .5))




```



## Race & Gender



```{r, eval = F}

analysis_data$oy_race_gender %>% 
  ggplot(aes(x = oy_flag, 
             y = percent)) + 
  geom_col(aes(fill = SEX_label)) + 
  facet_wrap(~race_ethnicity, 
             scales = 'free_y') + 
  scale_y_continuous(labels = scales::percent) + 
  geom_text(aes(label = scales::percent(percent, accuracy = .1), 
                group = SEX_label), 
            position = position_stack(vjust = .5), 
            size = 3) + 
  theme_classic() + 
  theme( strip.background = element_blank(), 
         # strip.text.x = element_text(face = 'bold'), 
         axis.title.x = element_blank()) + 
  scale_x_discrete(labels = function(x) str_wrap(x, 10)) + 
  labs(fill = "Gender")
  



```
```{r}


analysis_data$oy_race_gender_alternate %>% 
  # filter(SEX_label != "Female") %>% 
  ggplot(aes(x = race_alternate, 
             y = percent)) + 
  geom_col(aes(fill = SEX_label)) + 
  facet_wrap(~oy_flag) + 
  coord_flip() + 
  theme_classic() + 
  geom_text(aes(label = scales::percent(percent, accuracy = .1), 
                group = SEX_label), 
            size = 2, 
            position = position_stack(vjust= .5)) + 
  scale_y_continuous(labels = scales::percent) + 
  theme(
    strip.background = element_blank(), 
    strip.text.x = element_text(face = 'bold'), 
    axis.line.x = element_blank(), 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank()
  ) + 
  labs(
    x = "Race", 
    y = "Percentage", 
    fill = "Sex"
  ) + 
  scale_x_discrete(labels = function(x) str_wrap(x, 20)) + 
  scale_fill_manual(values = 
                      c(
                        "#fff7bc",
                        "#D5802B"
                      )) + 
  guides(
    # fill = FALSE
    )


```



```{r}
analysis_data$oy_race_gender_alternate %>% 
  filter(SEX_label != "Female") %>%
  ggplot(aes(x = race_alternate, 
             y = percent)) + 
  geom_col(aes(fill = SEX_label), 
           color = 'black') + 
  facet_wrap(~oy_flag) + 
  coord_flip() + 
  theme_classic() + 
  geom_text(aes(label = scales::percent(percent, accuracy = .1), 
                group = SEX_label), 
            size = 2, 
            position = position_stack(vjust= .5)) + 
  scale_y_continuous(labels = scales::percent) + 
  theme(
    strip.background = element_blank(), 
    strip.text.x = element_text(face = 'bold'), 
    axis.line.x = element_blank(), 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank()
  ) + 
  labs(
    x = "Race", 
    y = "Percentage and Race of ", 
    fill = "Sex"
  ) + 
  scale_x_discrete(labels = function(x) str_wrap(x, 20)) + 
  scale_fill_manual(values = 
                      c(
                        # "#fff7bc",
                        "#D5802B"
                      )) + 
  guides(
    fill = FALSE
    ) + 
  geom_errorbar(aes(ymin = percent, ymax = percent_upp), 
                width = .3)
```



```{r}

analysis_data$oy_race_gender_alternate %>% 
  filter(SEX_label != "Female") %>% 
  ggplot(aes(x = race_alternate, 
             y = percent)) + 
  geom_col(aes(fill = oy_flag), 
           position = 'dodge') + 
  scale_x_discrete(labels = function(x) str_wrap(x, 10))


```





## Population by PUMA

```{r, eval = T}

analysis_data$geo_puma_pop %>% 
  ggplot() + 
  geom_sf(aes(fill = opp_youth)) + 
  geom_sf_text(label = 
                  analysis_data$geo_puma_pop$opp_youth, 
               size = 1.8) + 
  theme_void() + 
  scale_fill_gradient(low = "#fff7bc",
                      high = "#D5802B",
                      na.value = "white") + 
  labs(fill = 
         "Opportunity Youth Population")



```

