---
title: "Opportunity Youth Household Structure (Preliminary)"
author: "Diego Rubi"
date: " `r format(Sys.Date(), '%B %d, %Y')`"
output: 
  powerpoint_presentation:
      reference_doc: reference_ppt.pptx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = F, 
                      message = F, 
                      fig.height = 5, 
                      fig.width = 8)
```

```{r load-libraries-functions}

source(
  here::here(
    'R', 
    'presentation_setup.R'
  )
)


```

```{r load-data}

analysis_data <- 
  load_data$load_population_estimates()

ipums_data <- 
  load_data$load_IPUMS_population_estimates()




```

```{r}
analysis_data$q1_demographics <- 
  
  map(analysis_data$q1_demographics, 
      .f =  ~
        .x %>% 
        modify_at(.x = ., 
                  .at = 'oy_flag', 
                  .f = figure_helpers$format_oy_variable))

analysis_data$q2_household <- 
  map(analysis_data$q2_household, 
      .f = ~ 
        .x %>%
              modify_at(.x = .,
                .at = 'oy_flag',
                .f = figure_helpers$format_oy_variable))

analysis_data[c("household_income_estimates", "household_type")] %<>%
  map(.f = ~ 
        .x %>% 
        
        mutate(oy_household = figure_helpers$format_oy_household(oy_household_full)))

analysis_data %<>%
  modify_at(
    .at = 'oy_flag', 
    .f = ~ figure_helpers$format_oy_variable(.)
  )

ipums_data %<>% 
  map(.f = ~ 
        .x %>% 
        mutate(oy_flag = figure_helpers$format_oy_variable(oy_flag)))


erase_y_axis = 
   theme(
    axis.text.y = element_blank(), 
    axis.line.y = element_blank(), 
    axis.title.y = element_blank(), 
    axis.ticks.y = element_blank()
  )

erase_x_legend = 
  theme(
    axis.title.x = element_blank()
  )


# Settings for the lower error bar
# other option is percent_low
error_bar_setting <- 
  rlang::sym('percent_low')

# color contrasts between Connected Youth and Opportunity Youth
# color_contrast <- 
#   c(`Connected Youth` = "#800000", `Opportunity Youth` = '#D5802B')

color_contrast <- 
  c(`Connected Youth` = "#767676", `Opportunity Youth` = '#D5802B')

color_contrast_alt <- color_contrast
names(color_contrast_alt) <- NULL

color_contrast_hh <- rep(color_contrast, 2)
names(color_contrast_hh) <- c("Connected Youth - HH", "Connected Youth - Non-HH", "Opportunity Youth - HH", "Opportunity Youth - Non-HH" )

color_palette <- 
  ggsci::pal_uchicago(palette = 'default')

color_palette_light <- 
  ggsci::pal_uchicago(palette = 'light')
  


```

```{r}
basic_plot <- function(data, fill_label = NULL){
  
  fill_col = names(data)[2] %>% rlang::sym()
  
  x_col =    names(data)[str_detect(names(data), "oy")] %>% rlang::sym()
  
  
  
  data %>% 
    filter(!str_detect(!!x_col, '(Everyone)|(Eveyrone)')) %>% 
    ggplot(aes(x = !!x_col, y = percent, fill = !!fill_col)) + 
    geom_col(color = 'black') + 
    geom_text(aes(label = if_else(percent > .05, 
                                  scales::percent(percent, accuracy = .1), 
                                  ""), 
                  group = !!fill_col), 
              position = position_stack(vjust = .5)) +
    theme_classic() + 
    scale_y_continuous(labels = function(x) scales::percent(x, accuracy = .1)) + 

    erase_y_axis + 
    erase_x_legend + 
    theme(axis.text.x = element_text(face = 'bold')) + 
    labs(
      caption = "Cells representing less than 5% suppressed for clarity") + 
    ggsci::scale_fill_uchicago()
    
    
}

single_plot <- function(data, x_col = NULL){
  
  if(is.null(x_col)){
    x_col =    names(data)[str_detect(names(data), "oy")] %>% rlang::sym()
  
  }else{
    x_col = rlang::sym(x_col)
  }
  
  data %>% 
    ggplot(aes(x = !!x_col, y = percent)) + 
    geom_col(color = 'black', 
             fill = color_contrast[2], 
             width = .7) + 
    geom_errorbar(aes(ymax = percent_upp, ymin = !!error_bar_setting), 
                  width = .2) + 
    theme_classic() + 
    # geom_text(aes(label = if_else(percent > .05, 
    #                               scales::percent(percent, accuracy = .1), 
    #                               ""))) + 
    
    erase_x_legend + 
    theme(axis.text.x = element_text(face = 'bold'))
}

wrap_x_labels = 
  scale_x_discrete(labels = function(x) str_wrap(x,10))

basic_plot_dodge <- function(data, fill_label = NULL, erase = FALSE){
  
  fill_col = names(data)[2] %>% rlang::sym()
  
  x_col =    names(data)[str_detect(names(data), "oy")] %>% rlang::sym()
  
  out <-  
    data %>% 
    filter(!str_detect(!!x_col, '(Everyone)|(Eveyrone)')) %>% 
    ggplot(aes(x = !!x_col, y = percent, fill = !!fill_col, group = !!fill_col)) + 
    geom_col(color = 'black', 
             position = 'dodge') + 
    # geom_text(aes(label = scales::percent(percent, accuracy = .1),
    #               group = !!fill_col), 
    #           position = position_dodge(width = .87), 
    #           vjust = 1.3) +
    # geom_errorbar(aes(ymin = percent, ymax = percent_upp, 
    #                   group = !!fill_col), 
    #               width = .5, 
    #               position = position_dodge(1))+
    theme_classic() + 
    scale_y_continuous(labels = function(x) scales::percent(x, accuracy = .1)) + 
    theme(axis.text.x = element_text(face = 'bold')) +
    erase_x_legend
  
  if(erase == TRUE){
    out <-  out + erase_y_axis
  }else{
    out <-  out
  }
    
    # erase_y_axis + 
    # erase_x_legend + 
    # theme(axis.text.x = element_text(face = 'bold'))
    # scale_fill_manual(values = color_contrast_hh)
  
  return(out)
}


```

### Data Sources:

- [IPUMS USA](https://usa.ipums.org/usa/)
    - IPUMS at the University of Minnesota
    - IPUMS pre-processes PUMS files to enable / facilitate research 
        - Example: NOC (U.S. Census) vs. NCHILD (IPUMS) are at household level and individual level respectively
- [U.S. Census Bureau](https://www.census.gov/programs-surveys/acs/microdata.html)
    - Raw PUMS files directly from the U.S. Census Bureau

### Research Questions of Interest

  - What is each household type, and what is their composition?
  - What is the average household size for each type of household?
  - What is average household income by household type?
  - What is an Opportunity Youth's average total personal income?
  - What percentage of households are food stamp recipients?
  - How many youth identify as the reference person?
  - How many youth have at least one child?
  - How many youth have multiple children?
  - What is the household median income for each household type?
  - What is the median income by Chicago region for reference persons?
  - What are youth's average total personal income? 
  
### Research Questions (cont.)

  - How many youth are married?
  - How many youth are partnered?
  - How many households are foodstamp recipients?
  - What type of households do
  - What percentage of each household has broadband access?
  - What percentage of each household has computer access?
  - Have Opportunity Youth moved within the last year?
  - What are Connected Youths' primary form of transportation?
  - What are Connected Youths' average commute time?

# Refresher

### How are Opportunity Youth defined?

- _Opportunity Youth:_ Individuals that are 16 to 24 years old who are simultaneously unemployed/ not in the labor force and not enrolled in school (disconnection)
- _Connected Youth:_ Individuals that are 16 to 24 years old who are either employed, enrolled in school, or both.
- _Everyone Else:_ Individuals outside of this age category who may be enrolled in school or employed




### 14% of 16 - 24 year olds are Opportunity Youth

```{r}
ss <- analysis_data$q1_demographics$total_oy_population %>% 
  select(oy_flag, n) %>% 
  spread(oy_flag, n) %>% 
  unlist()

analysis_data$q1_demographics$total_oy_population %>% 
  single_plot() + 
  scale_x_discrete(labels = function(x) paste0(x, figure_helpers$fig_help$newline, 
                                               figure_helpers$fig_help$N, 
                                               ss %>% format(big.mark = ','))) + 
  labs(y = "Percent", 
       caption = paste0(figure_helpers$citations['PUMS'], 
                        figure_helpers$fig_help$newline)) + 
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  geom_text(aes(label = scales::percent(percent, accurary = .01)), 
            position = position_stack(vjust = .5))


```


# Results

### How are households categorized?

```{r}

# **Household-level:**
# 
# - *OY Household:* There is at least one 'Opportunity Youth' in the household but no 'Connected Youth' in the household. There may be 'Everyone Else' in the household.
# - *OY & CY Household:* There are both 'Opportunity Youth' and 'Connected Youth' in the household and possibly 'Everyone Else'.
# - *CY Only Household:* There are only 'Connected Youth' and no 'Opportunity Youth' or 'Everyone Else' in the household.

present_func <- function(var){
  
  var <- 
    if_else(!is.na(var), 
            "Present",
            "Not Present")
  
  return(var)
  
}

analysis_data$household_chart %>% 
  mutate(across(c(connected_youth, everyone_else, opp_youth), 
                .fns = ~ present_func(.))) %>% 
  select(c(1,3,2,4)) %>% 
  pretty_flex() %>% 
  flextable::set_header_labels(
    oy_household_full = "Household Type", 
    everyone_else = "Everyone Else", 
    connected_youth = "Connected Youth", 
    opp_youth = "Opportunity Youth"
  )
  
```


### Opportunity Youth households are more likely to contain 'Everyone Else'

```{r}

analysis_data$hh_composition %>%
  ggplot(aes(x = oy_household_full, y = percent, fill = oy_flag, group = oy_flag)) + 
  geom_col(position = 'dodge', 
           color = 'black') + 
  
    geom_errorbar(aes(ymin = !!error_bar_setting, ymax = percent_upp), 
            position = position_dodge(width = .95), 
            width = .2) + 
  
  geom_label(aes(
    label = scales::percent(percent, accuracy = .1)
                    # y = percent_upp + .03
    ), 
    position = position_dodge(width = .9), 
    size = 2.7, 
    color = 'black', 
    fill = 'white') + 
  

  theme_classic() + 
  
  labs(y = "Percent", 
       x = "Type of Household", 
       fill = "OY Type") + 
  wrap_x_labels + 
  
  ggsci::scale_fill_uchicago() + 
  scale_y_continuous(labels = scales::percent)

```



```{r, eval = F}

### Household Composition: Alternate
change_zero <- function(vector){
  
  vector <- 
    if_else(is.na(vector), 0, 
            vector)
  
  return(vector)
  
}



analysis_data$household_chart_percent %>%
  mutate(across(is.double, ~ change_zero(.))) %>% 
  select(1, 3, 2, 4) %>% 
  ungroup() %>% 
  mutate(across(is.double,.fns = ~ scales::percent(., accuracy = .1))) %>% 
  pretty_flex() %>%
  flextable::set_header_labels(
    oy_household_full = "Household Type", 
    everyone_else = "Everyone Else", 
    connected_youth = "Connected Youth", 
    opp_youth = "Opportunity Youth"
  )


```





### Less than 3% of households contain only Opportunity Youth

```{r, fig.width = 9}
ss <- 
  analysis_data$household_type %>% 
  select(oy_household_full, n) %>%
  mutate(label = paste0(oy_household_full, "\n", "N = ", n %>% format(big.mark = ','))) %>%
  select(oy_household_full, label) %>% 
  spread(oy_household_full, label) %>% 
  unlist()

analysis_data$household_type %>% 
  # filter(oy_household_full != "EE Only Household") %>% 
  # mutate(oy_household = figure_helpers$format_oy_household(oy_household)) %>% 
  single_plot(x_col = 'oy_household_full') +
  geom_label(aes(label = scales::percent(percent, accuracy = .1))) +
  scale_y_continuous(labels = scales::percent, breaks = c(0, .2,   .4,  .6,  .8)) + 
  labs(y = "Percent", 
       caption = paste0(figure_helpers$citations['PUMS'], 
                        figure_helpers$fig_help$newline, 
                        figure_helpers$suppress), 
       x = "Household Type") + 
  scale_x_discrete(labels = ss) + 
  theme(axis.text.x = element_text(size = 8)) 
```





### Both Connected Youth and Opportunity Youth live in larger households 

```{r}

analysis_data$num_people_lived_with %>% 
  ggplot(aes(x = oy_household_full, y = percent)) + 
  geom_col(color = 'black', 
           fill = color_contrast[2]) +
  # geom_point(aes(size = percent_upp, color = oy_household)) +
  
  geom_errorbar(aes(ymin = !!error_bar_setting, ymax = percent_upp), 
                width = .2) +
  geom_label(aes(label = format(percent, digits = 2))) +
  # geom_segment(aes(x = oy_household, xend = oy_household, y = 0,yend = percent)) + 
  theme_classic() + 
  wrap_x_labels + 
  labs(y = "Average number of people in household", 
       x = "Type of Household", 
       size = "Standard Error") + 
  ggsci::scale_color_uchicago()


```






### Opportunity Youth are half as likely as Connected Youth to be the reference person


```{r}
ss <- analysis_data$q2_household$head_of_household_alt %>% group_by(oy_hh_flag) %>% summarize(n = sum(n)) %>% spread(oy_hh_flag, n) %>% unlist()


analysis_data$q2_household$head_of_household_alt %>%
  filter(!str_detect(oy_hh_flag, "Eve")) %>% 
  single_plot(x_col = 'oy_hh_flag') + 
  wrap_x_labels + 
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(caption = figure_helpers$citations['PUMS'], 
       y = "Percent") + 
  scale_x_discrete(labels = function(x) paste0(x, 
                                               figure_helpers$fig_help$newline, 
                                               figure_helpers$fig_help$N,
                                               (ss[-c(3:4)]) %>% format(big.mark = ','))) + 
  geom_label(aes(label = scales::percent(percent)))
```


### Opportunity Youth are three times as likely to have at least one child

```{r, fig.height = 6}
# children are now individual level so I can say, does a given individual have at least one child?

ipums_data$at_least_one_child %>% 
  mutate(plus_minus = 
           paste0(scales::percent(percent, suffix = "", accuracy = .1),
                  " (+/-",
                  (2*percent_se) %>% scales::percent(suffix = "", accuracy = .1), 
                  ")")) %>% 
  # ggplot(aes(x = oy_flag, y = percent, group = at_least_one_child, fill = at_least_one_child)) + 
  # geom_point() +
  basic_plot_dodge(erase = F) +
  geom_errorbar(aes(ymin = !!error_bar_setting, ymax = percent_upp,
                    group = at_least_one_child),
                width = .2,
                position = position_dodge(.9))+
  labs(fill = "Parental Status") + 
  # geom_text(
  #   aes(label = scales::percent(percent), 
  #       group = at_least_one_child), 
  #   position = position_dodge(width = .9), 
  #   vjust = 1.2, 
  #   size = 3.8
  # ) + 
  wrap_x_labels + 
  scale_fill_manual(values = color_contrast_alt) + 
  labs(
    caption = paste0(figure_helpers$citations['IPUMS'],
                     figure_helpers$fig_help$newline,
                     figure_helpers$suppress),
    y = "Percent"
  ) + 
  geom_label(
    aes(label =
          # if_else(percent >= 0,
          #         scales::percent(percent, accuracy = .1),
          #         ""),
          scales::percent(percent, suffix = "", accuracy = .1),
        # y = percent_upp + .02,
        group = at_least_one_child),
    position = position_dodge(width = .9),
    # size = 4,
    color = 'black', 
    fill = 'white'
  )
  



```

### Opportunity Youth are more likely to have multiple children

```{r, eval = T}

# analysis_data$hh_num_children %>% 
#   filter(!str_detect(oy_hh_flag, 'Eve')) %>% 
#   
#   ggplot(aes(x = oy_hh_flag, y = percent)) + 
#   geom_col(color = 'black', 
#            fill = color_contrast[2]) + 
#   geom_errorbar(aes(ymax = percent_upp, ymin = !!error_bar_setting), 
#                 width = .2) + 
#   theme_classic() + 
#   geom_text(aes(label = round(percent, digits = 2)), 
#             position = position_stack(vjust = .5)) + 
#   # labs( y = "Average Number of Children") + 
#   erase_x_legend + 
#   theme(axis.text.x = element_text(face = 'bold')) + 
#   labs(caption = "Cells representing less than 5% suppressed for clarity", 
#        y = 'Average number of children') + 
#   wrap_x_labels 
  
ipums_data$percent_children %>% 
  filter(oy_flag != "everyone_else") %>% 
  basic_plot_dodge(erase = FALSE) + 
  geom_errorbar(aes(ymin = !!error_bar_setting, ymax = percent_upp,
                    group = NCHILD),
                width = .2,
                position = position_dodge(.9)) + 
  # geom_text(aes(label = scales::percent(percent, accuracy = .1), 
  #               group = NCHILD), 
  #           position = position_dodge(width = .9)) + 
  labs(
    caption = paste0(figure_helpers$citations['IPUMS'], 
                     figure_helpers$fig_help$newline, 
                     figure_helpers$suppress), 
    fill = "Number of Children", 
    y = "Percent"
  ) + 
  geom_label(
    aes(label = 
          if_else(percent > .05, 
                  scales::percent(percent, accuracy = .1),
                  ""),
        group = NCHILD), 
    position = position_dodge(width = .9), 
    size = 2.5, 
    color = 'black', 
    fill = 'white'
    # vjust = 1.4
  ) + 
  ggsci::scale_fill_uchicago(palette = 'default')



```





### The median income of OY Only Households is less than half that of other households 

```{r, fig.height = 6}
  

analysis_data$household_type_median_income %>% 
  ggplot(aes(x = fct_reorder(oy_household_full, percent), y = percent)) + 
  geom_col(color = 'black', 
           fill = color_contrast[2]) + 
  geom_errorbar(aes(ymax = percent_upp, ymin = !!error_bar_setting), 
                width = .2) + 
  geom_hline(yintercept = analysis_data$chicago_household_median_income$percent, 
             linetype = 'dotted') + 
  theme_classic() + 
  geom_label(aes(label = scales::dollar(percent)), 
             color = 'black', 
             fill = 'white', 
             size = 2.5) + 
  # labs( y = "Average Number of Children") + 
  erase_x_legend + 
  theme(axis.text.x = element_text(face = 'bold')) + 
  labs(
    caption = figure_helpers$citations['PUMS'], 
    y = "Median Household Income"
    ) + 
  scale_y_continuous(labels = scales::dollar) + 
  wrap_x_labels + 
  annotate(geom = 'text', 
           y = analysis_data$chicago_household_median_income$percent + 1000,
           x = 1,
           label = scales::dollar(analysis_data$chicago_household_median_income$percent))



```


### Median Income OY Reference Person by PUMA

```{r}

analysis_data$median_income_hh_by_puma_region %>% 
  filter(str_detect(oy_hh_flag, "(Connected)")) %>% 
  ggplot(aes(x = PUMA_region, y = n)) + 
  geom_col(aes(fill = oy_hh_flag), 
           position = 'dodge', 
           color = 'black') + 
  geom_errorbar(aes(ymin = n_low, ymax = n_upp, group = oy_hh_flag), 
                width = .2, 
                linetype = 'dashed', 
                position = position_dodge(width = .9)) +
  geom_label(aes(label = scales::dollar(n), 
                group = oy_hh_flag), 
            position = position_dodge(width = .9), 
            color = 'black', 
            fill = 'white', 
            size = 2.5) +

  scale_y_continuous(labels = scales::dollar) + 
  theme_classic() + 
  labs(y = "Median Income", 
       x = "PUMA Region", 
       fill = "Head of Household Status") + 
  theme(
    strip.background = element_blank(), 
    strip.text = element_text(face = 'bold')
  ) + 
  ggsci::scale_fill_uchicago()



```




### Opportunity Youth have 1/4th the total average personal income of Connected Youth


```{r}

analysis_data$q2_household$average_income %>% 
  # filter(!(str_detect(oy_flag, "Eve"))) %>%
  ggplot(aes(x = oy_flag, y = percent)) + 
  geom_col(color = 'black', 
           fill = color_contrast[2]) + 
  geom_errorbar(aes(ymax = percent_upp, ymin = !!error_bar_setting), 
                width = .2)+
  theme_classic() + 
  geom_text(aes(label = scales::dollar(percent)), 
            position = position_stack(vjust = .5), 
            size = 3.7) + 
  # labs( y = "Average Number of Children") + 
  erase_x_legend  +
  theme(axis.text.x = element_text(face = 'bold')) + 
  labs(
    y = "Average Total Personal Income",
    caption = figure_helpers$citations['PUMS']
  ) +
  wrap_x_labels + 
  scale_y_continuous(labels = scales::dollar, limits = c(0,40000)) 


  


```




```{r, eval = F}

# Income Characteristics: Income Bracket

analysis_data$hh_income_bracket %>% 
  # filter(oy_flag != "Everyone Else") %>% 
  basic_plot() + 
  labs(fill = "Income Bracket") + 
  wrap_x_labels + 
  scale_fill_manual(values = c(color_contrast_alt[2],color_palette(9), color_palette_light(9)))


```



### Opportunity Youth are unlikely to be in a relationship
```{r, fig.width = 10}

# person level
p1 <- 
  analysis_data$q2_household$married %>% 
  basic_plot_dodge(erase = F) + 
  labs(fill = "Marital Status", 
       caption = paste0(figure_helpers$citations['PUMS'], 
                        figure_helpers$fig_help$newline, 
                        figure_helpers$suppress), 
       y = "Percent") + 
  wrap_x_labels  + 
  scale_fill_manual(values = c(color_contrast_alt[2], rev(color_palette(5)))) + 
  geom_errorbar(aes(ymin = !!error_bar_setting, ymax = percent_upp,
                    group = MAR_label),
                width = .2,
                position = position_dodge(.9)) + 
  geom_label(
    aes(label = 
          if_else(percent > .05, 
                  scales::percent(percent, accuracy = .1),
                  NULL),
        group = MAR_label), 
    position = position_dodge(width = .9), 
    size = 2.5, 
    color = 'black', 
    fill = "white"
  )

# person level
p2 <- 
  analysis_data$q2_household$partnered %>% 
  basic_plot_dodge(erase = F) + 
  geom_errorbar(aes(ymin = !!error_bar_setting, ymax = percent_upp,
                    group = PARTNER_label),
                width = .2,
                position = position_dodge(.9)) + 
  labs(fill = "Unmarried partner household", 
       caption = paste0(figure_helpers$citations['PUMS'], 
                        figure_helpers$fig_help$newline, 
                        figure_helpers$suppress), 
       y = "Percent") + 
  wrap_x_labels  + 
  ggsci::scale_fill_uchicago() + 
  geom_label(
    aes(label = 
          if_else(percent > .05, 
                  scales::percent(percent, accuracy = .1),
                  NULL),
        group = PARTNER_label), 
    position = position_dodge(width = .9), 
    size = 2.5, 
    color = 'black', 
    fill = 'white'
  )
  

p1



```


```{r}
p2
```


### Opportunity Youth are more likely to have lived in the same house the prior year

```{r}

# person level
analysis_data$q2_household$migration %>% 
  basic_plot_dodge(erase = F) + 
  labs(fill = "Mobility status (lived here 1 year ago)", 
       caption = paste0(figure_helpers$citations['PUMS'], 
                        figure_helpers$fig_help$newline,
                        figure_helpers$suppress), 
       y = "Percent") + 
  wrap_x_labels + 
  scale_fill_manual(values = c(color_palette(3), color_contrast_alt[2])) + 
  geom_errorbar(aes(ymin = !!error_bar_setting, ymax = percent_upp), 
                position = position_dodge(width = .9),
                width = .2) + 
  geom_label(
    aes(label = 
          if_else(percent > .05, 
                  scales::percent(percent, accuracy = .1),
                  NULL),
        group = MIG_label), 
    position = position_dodge(width = .9), 
    size = 2.5, 
    color = 'black', 
    fill = 'white'
  )

```

### Mobility: Mode of Transportation for Connected Youth
```{r}
# Person level
analysis_data$q2_household$transportation_mode %>% 
  filter(oy_flag == "Connected Youth") %>% 
  mutate(JWTR_label = str_remove(JWTR_label, "\\(.*\\)"), 
         JWTR_label = fct_reorder(JWTR_label, n)) %>% 
  basic_plot_dodge(erase = F) + 
  labs(fill = "Means of transportation to work", 
       caption = paste0(figure_helpers$citations['PUMS'], 
                        figure_helpers$fig_help$newline, 
                        figure_helpers$suppress),
       y = "Percent") + 
  wrap_x_labels + 
  geom_errorbar(aes(ymin = percent, ymax = percent_upp), 
                position = position_dodge(width = .9),
                width = .2) + 
  scale_fill_manual(values = c(color_palette(9), color_palette_light(9))) + 
  geom_label(aes(label = if_else(percent > .05, 
                                scales::percent(percent, accuracy = .1), 
                                NULL)), 
            position = position_dodge(width = .9), 
            
            size = 2.5, 
            color = 'black', 
            fill = 'white')

```

### Most Connected Youth commute less than ten minutes to work

```{r}

# mode of transportation
# person level 
analysis_data$q2_household$commute_bracket %>% 
  filter(oy_flag != "Opportunity Youth") %>% 
  mutate(commute_bracket = forcats::fct_relevel(commute_bracket, "100 +", after = 11)) %>% 
  basic_plot_dodge(erase = F) +
  labs(fill = "Commute Time (minutes)", 
       caption = figure_helpers$citations['PUMS'], 
       y = "Percent") + 
  wrap_x_labels + 
  geom_errorbar(aes(ymin = percent, ymax = percent_upp),
                position = position_dodge(width = .9),
                width = .2) + 
  scale_fill_viridis(discrete = TRUE)



```


### OY Only Households are twice as likely to receive foodstamps

```{r, eval = T}
# household level
analysis_data$foodstamps_per_household %>% 
  basic_plot_dodge() + 
  geom_errorbar(aes(ymin = !!error_bar_setting, ymax = percent_upp), 
                position = position_dodge(width = .9), 
                width = .2) +
  geom_label(aes(label = scales::percent(percent, accuracy = .1)), 
            position = position_dodge(width = 1), 
            color = 'black', 
            fill = 'white', 
            size = 2.5)+ 
  labs(fill = "Food Stamp Recipient", 
       caption = figure_helpers$citations['PUMS'], 
       y = 'Percent') +
  wrap_x_labels + 
  scale_fill_manual(values = color_contrast_alt) 



```



### Household Type

```{r, eval = T}
# household level
analysis_data$hht %>% 
  filter(oy_household_full != "All Other Households") %>%
  mutate(HHT_label = fct_reorder(HHT_label, percent)) %>% 
  
  basic_plot_dodge() + 
  geom_errorbar(aes(ymin = !!error_bar_setting, ymax = percent_upp, group = HHT_label), 
                position = position_dodge(width = .9), 
                width = .2) +
  geom_label(aes(label =if_else(percent > .05, 
                               scales::percent(percent, suffix = "", accuracy = .1), 
                               NULL),
                group = HHT_label), 
            size = 2.5, 
            color = 'black', 
            fill = 'white',
            position = position_dodge(width = .9)) +
  labs(fill = "Type of Household", 
       y = 'Percent') + 
  scale_fill_manual(values = c(color_contrast[2], color_palette(8)), 
                    labels = function(x) str_wrap(x,30)) + 
  theme(
    legend.position = 'right')

  



```


### OY Only Households are less likely to have access to broadband internet

```{r}

analysis_data$HISPEED_internet %>% 
  basic_plot_dodge() + 
  geom_errorbar(aes(ymin = !!error_bar_setting, ymax = percent_upp, group = HISPEED_label), 
                width = .2, 
                
                position = position_dodge(width = .9)) + 
  geom_label(aes(label = scales::percent(percent, suffix = "", accuracy = .1), 
                group = HISPEED_label), 
            position = position_dodge(width = .9), 
            color = 'black', 
            fill = 'white', 
            size = 2.5) + 
  ggsci::scale_fill_uchicago() + 
  labs(
    y = "Percent", 
    fill = "Broadband Access"
  )

```


### OY Only Households are less likely to have laptop access

```{r}
analysis_data$LAPTOP %>%
  filter(percent != 0) %>% 
  basic_plot_dodge() + 
  geom_errorbar(aes(ymin = !!error_bar_setting, ymax = percent_upp, 
                    group = LAPTOP_label), 
                width = .2, 
                position = position_dodge(width = .9)) + 
  geom_label(aes(label = 
                   scales::percent(percent, accuracy = .1, suffix = "")), 
                 color = 'black', 
                 fill = 'white', 
                 size = 2.5, 
            position = position_dodge(width = .9)) + 
  ggsci::scale_fill_uchicago() + 
  labs(
    y = "Percent", 
    fill = "Laptop Availability"
  )


```



```{r, eval = F}

### Household Characteristics: Workers in Family (Pending)
# household level
analysis_data$workers_in_fam %>% 
  basic_plot_dodge() + 
  labs(fill = "Number of workers in Family") + 
  wrap_x_labels + 
  scale_fill_manual(values = c(color_contrast_alt[2], color_palette_light(4)))
  

```



```{r, eval = F}

### Same-sex households (Pending)
# household level
analysis_data$same_sex_married %>% 
  basic_plot_dodge() + 
  labs(fill = "Same-Sex Married Household") + 
  wrap_x_labels

```



```{r, eval = F}

### Multi-generational Families (Pending)
# household level
analysis_data$multigen %>% 
  basic_plot_dodge() + 
  labs(fill = "Multigenerational Family") + 
  wrap_x_labels




```








# Questions?

# Thank you!


